---
title: "Linear Discriminant Analysis"
output: github_document
---

```{r setup, include=TRUE}
library(datasets)
library(MASS)
library(ggplot2)
library(dplyr)
```

## Linear Discriminant Analysis

Linear Discriminant Analysis (LDA) is a dimensionality reduction technique. As the name implies dimensionality reduction techniques reduce the number of dimensions (i.e. variables) in a dataset while retaining as much information as possible. Linear Discriminant Analysis, or LDA, uses the information from features to create a new axis and projects the data on to the new axis in such a way as to minimizes the variance and maximizes the distance between the means of the  classes.


```{r lda}
summary(iris)
lda.fit<-lda(Species~.,data=iris)
```

Performing Linear Discriminant Analysis on Iris dataset for classification per Species

```{r lda2}
lda.fit
plot(lda.fit)

predictions <- lda.fit %>% predict(iris)
iris <- cbind(iris, predict(lda.fit)$x)
ggplot(iris, aes(LD1, LD2)) +
  geom_point(aes(color = Species))
```

Reference:
https://towardsdatascience.com/linear-discriminant-analysis-in-python-76b8b17817c2#:~:text=Linear%20Discriminant%20Analysis%20(LDA)%20is,as%20much%20information%20as%20possible.

https://www.displayr.com/linear-discriminant-analysis-in-r-an-introduction/
